---
title: "smb_fitness_extra_analyses"
author: "Joe Gunn"
date: "2023-07-10"
output: html_document
---
### STEP 1: Define the von Bertlanffy growth model; run the Rmd chunk below.
In this step, we use the vbFuns() function in the R package FSA to generate a function using the "typical" parameterization of the von Bertlanffy equation.

##### Define von Bertalanffy growth model:
```{r}
vb <- vbFuns(param = "Typical")
```

### STEP 2: Parameterize the von Bertalanffy model for study groups of interest.

#### 2a: Generate global growth model.
In this step, we parameterize the von Bertalanffy growth model for the global dataset (all samples), including all fish in the study (across Big Sugar Creek and Elk River, both sexes and all ancestry groups).

##### 2a.1. Load full back-calculation data and generate dataset; run the Rmd chunk below.

##### Separate global dataset by river:
```{r}
# Load full bc data
load("data/bc_data/full_bc_data.Rda")

# Get global samples (all samples)
full_bc_data_global <- full_bc_data

# save Big Sugar Creek data for downstream analyses
save(full_bc_data_global, file = "data/global_data/full_bc_data_global.Rda")
```

##### 2a.2. Generate model starting values for the global dataset; run the Rmd chunk below.

##### load data and generate starting values
```{r}
# Load full bc data
load("data/bc_data/full_bc_data.Rda")

# Generate starting values for optimizing the von Bertalanffy model
global_starts <- vbStarts(bc_tl ~ annulus, 
                          data = full_bc_data)
```

<b>Starting values used for global model</b>: <br>

<b>L<sub>inf</sub></b>: 382.0127 <br>
<b>K</b>: 0.4094681 <br>
<b>t<sub>0</sub></b>: -0.07894399 <br>

##### 2a.3. Model growth for the global dataset and extract model coefficients; run the Rmd chunk below.
In this step, we run the non-linear von Bertalanffy model for the global dataset and extract model coefficients from the summary table given with the summary() function.

##### Model von Bertalanffy growth:
```{r}
# Run von Bertalanffy growth model using back-calculated total length, modeled by annulus number (consensus age)
global <- nls(bc_tl ~ vb(annulus,
                         Linf,
                         K,
                         t0), 
              data = full_bc_data, 
              start = global_starts)

# Extract global data coefficients
global_coefficients <- summary(global)
```

<b>Parameter estimates for global model</b>: <br>

<b>L<sub>inf</sub></b>: 440.94667 <br>
<b>K</b>: 0.28338 <br>
<b>t<sub>0</sub></b>: -0.30329 <br>

Parameter estimates are given in Table S3a in the final manuscript.

##### 2a.4. Perform bootstrapping of global data to estimate 95% confidence intervals around each parameter estimate; run the Rmd chunk below.
In this step, we perform bootstrapping on the global dataset with 9999 bootstrap replicates and calculate 95% confidence intervals around the parameter estimates given in the global model (see Step 2a.3).

##### Perform bootstrapping on global model and plot histogram of results: `figures/bootstrap_plots/global.pdf`
```{r}
# Conduct bootstrapping on parameter estimates
set.seed(34349) # Set seed so that results are reproducible
global_bootstrap <- Boot(global, 
                         R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates
global_confint <- confint(global_bootstrap, 
                          level = 0.95)

# Save global bootstrap data for downstream analysis
save(global_bootstrap, file = "data/global_data/global_bootstrap.Rda")
```

<b>Confidence intervals for global parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 408.4898281 - 485.3221992 <br>
<b>K</b>: 0.2287383 - 0.3429249 <br>
<b>t<sub>0</sub></b>: -0.5030147 - -0.1412999 <br>

Results from bootstrapping and confidence intervals are given in Table S3a in the final manuscript.

##### 2a.5. Plot histograms of bootstrap results for global data; run the Rmd chunk below.
In this step, we plot histograms of the bootstrap replicate results to visualize 95% confidence intervals.

##### Plot histograms of bootstrap results for the global data: `figures/bootstrap_plots/global.pdf`
```{r}
# Load global bootstrap data
load("data/global_data/global_bootstrap.Rda")

# Convert bootstrap data to dataframe
global_boot_data <- data.frame(global_bootstrap$t)

# Plot histogram of Linf bootstrap values
Linf <- ggplot() + 
  geom_histogram(data = global_boot_data, aes(x = Linf, fill = "grey"), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 408.4898281, xend = 485.3221992), y = 1300, yend = 1300, color = "black", stat = "identity") +
  geom_point(aes(x = 440.94667, y = 1300, fill = "grey"), color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "L∞", y = "Frequency") +
  scale_fill_manual(values = c("grey")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(L[infinity])))

# Plot histogram of K bootstrap values
K <- ggplot() + 
  geom_histogram(data = global_boot_data, aes(x = K, fill = "grey"), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 0.2287383, xend = 0.3429249), y = 1050, yend = 1050, color = "black", stat = "identity") +
  geom_point(aes(x = 0.28338, y = 1050, fill = "grey"), color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "K", y = "Frequency") +
  scale_fill_manual(values = c("grey")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.title.x = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(K[B])))

# Plot histogram of t<sub>0</sub> bootstrap values
t0 <- ggplot() + 
  geom_histogram(data = global_boot_data, aes(x = t0, fill = "grey"), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = -0.5030147, xend = -0.1412999), y = 1050, yend = 1050, color = "black", stat = "identity") +
  geom_point(aes(x = -0.30329, y = 1050, fill = "grey"), color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "t0", y = "Frequency") +
  scale_fill_manual(values = c("grey")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic("t"[0]))) 

# Plot histograms of each parameter estimate together for the global data
pdf("figures/bootstrap_plots/global.pdf", width = 15, height = 4)

plot_grid(Linf, 
          K, 
          t0, 
          nrow = 1, 
          ncol = 3)

dev.off()
```

##### 2a.5. Perform bootstrapping of predicted average length-at-age values and calculate 95% confidence intervals; run the Rmd chunk below.
In this step, we perform bootstrapping over average length-at-age estimates predicted for the dataset and calculate 95% confidence intervals around the parameter estimates.

##### Conduct bootstrapping over predicted values for the global dataset:
```{r}
# Define function to predict average length-at-age values for each age (to plot the best fit curve for the von Bertlanffy growth model)
predict_function <- function(x) {
  predict(x, 
          data.frame(annulus = annuli))
}

# Define range of theoretical annulus values over which to calculate length-at-age
annuli <- seq(-1, 15, by = 0.2)

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
global_iboot <- Boot(global, 
                     f = predict_function,
                     R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
global_iconfint <- confint(global_iboot, 
                           level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
global_predict <- data.frame(annuli,
                             predict(global, data.frame(annulus = annuli)),
                             global_iconfint)

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(global_predict) <- c("annulus","fit","lci","uci")

# Save the predicted global model dataset for donwstream analyses
save(global_predict, file = "data/global_data/global_predict.Rda")
```

##### 2a.6. Plot von Bertlanffy growth for the global dataset; run the Rmd chunk below.

##### Plot the von Bertalanffy model for the global dataset: `figures/vb_plots/global.pdf`
```{r}
# Load in global predicted dataset to plot best fit curve
load("data/global_data/global_predict.Rda")

# Load in full back-calculuated data to plot all data points
load("data/global_data/full_bc_data_global.Rda")

# Plot von Bertalanffy model
pdf("figures/global.pdf", width = 5, height = 4)

ggplot() + 
  geom_ribbon(data = global_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = "grey", 0.3), show.legend = F) +
  geom_point(data = full_bc_data_global, aes(x = annulus, y = bc_tl, fill = "grey"), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.2)) + 
  geom_line(data = global_predict, aes(x = annulus, y = fit)) +
  geom_hline(yintercept = 440.94667, linetype = "solid", size = 0.5) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("grey")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,500), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

This figure is the basis for Figure 2a in the final manuscript.

#### 2b: Generate growth model for each river (Big Sugar Creek, Elk River) separately.
In this step, we parameterize the von Bertalanffy growth model for each stream population (Big Sugar Creek and Elk River) to assess potential differences in growth due to stream ecology.

##### 2b.1. Load full back-calculation data and separate into datasets by stream; run the Rmd chunk below.

##### Separate global dataset by river:
```{r}
# Load full bc data
load("data/bc_data/full_bc_data.Rda")

# Get Big Sugar Creek samples only
full_bc_data_bs <- full_bc_data %>%
  filter(river == "Big_Sugar")

# Get Elk River samples only
full_bc_data_elk <- full_bc_data %>%
  filter(river == "Elk_River")

# save Big Sugar Creek data for downstream analyses
save(full_bc_data_bs, file = "data/river_data/full_bc_data_bs.Rda")

# save Elk River data for downstream analyses
save(full_bc_data_elk, file = "data/river_data/full_bc_data_elk.Rda")
```

##### 2b.2. Generate model starting values for each river; run the Rmd chunk below.

##### load data and generate starting values for rivers:
```{r}
# Load full bc data for Big Sugar Creek
load("data/river_data/full_bc_data_bs.Rda")

# Load full bc data for Elk River
load("data/river_data/full_bc_data_elk.Rda")

# Generate starting values for optimizing the von Bertalanffy model in Big Sugar Creek
bs_starts <- vbStarts(bc_tl ~ annulus, 
                      data = full_bc_data_bs)

# Generate starting values for optimizing the von Bertalanffy model in Elk River
elk_starts <- vbStarts(bc_tl ~ annulus, 
                       data = full_bc_data_elk)
```

<b>Starting values used for Big Sugar Creek model</b>: <br>

<b>L<sub>inf</sub></b>: 455.5685 <br>
<b>K</b>: 0.2670832 <br>
<b>t<sub>0</sub></b>: -0.2545336 <br>

<b>Starting values used for Elk River model</b>: <br>

<b>L<sub>inf</sub></b>: 381.0677 <br>
<b>K</b>: 0.4223898 <br>
<b>t<sub>0</sub></b>: -0.09427 <br>

##### 2b.3. Model growth for each river population and extract model coefficients; run the Rmd chunk below.
In this step, we run the non-linear von Bertalanffy model for each river dataset and extract model coefficients from the summary table given with the summary() function.

##### Model von Bertalanffy growth for rivers:
```{r}
# Run von Bertalanffy growth model using back-calculated total length, modeled by annulus number (consensus age)
bs <- nls(bc_tl ~ vb(annulus,
                     Linf,
                     K,
                     t0), 
          data = full_bc_data_bs, 
          start = bs_starts)

# Extract Big Sugar data coefficients
bs_coefficients <- summary(bs)

# Run von Bertalanffy growth model using back-calculated total length, modeled by annulus number (consensus age)
elk <- nls(bc_tl ~ vb(annulus,
                      Linf,
                      K,
                      t0), 
           data = full_bc_data_elk, 
           start = elk_starts)

# Extract Big Sugar data coefficients
elk_coefficients <- summary(elk)


```

<b>Parameter estimates for Big Sugar Creek model</b>: <br>

<b>L<sub>inf</sub></b>: 464.18395 <br>
<b>K</b>: 0.25180 <br>
<b>t<sub>0</sub></b>: -0.30943 <br>

<b>Parameter estimates for Elk River model</b>: <br>

<b>L<sub>inf</sub></b>: 430.14416 <br>
<b>K</b>: 0.30578 <br>
<b>t<sub>0</sub></b>: -0.28753 <br>

Parameter estimates are given in Table S3b in the final manuscript.

##### 2b.4. Perform bootstrapping of river population datasets to estimate 95% confidence intervals around each parameter estimate; run the Rmd chunk below.
In this step, we perform bootstrapping on each river population dataset with 9999 bootstrap replicates and calculate 95% confidence intervals around the parameter estimates given in the Big Sugar Creek and Elk River models (see Step 2b.3).

##### Perform bootstrapping on Big Sugar Creek and Elk River models:
```{r}
# Conduct bootstrapping on parameter estimates for Big Sugar Creek model
set.seed(49504) # Set seed so that results are reproducible
bs_bootstrap <- Boot(bs, 
                     R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for Big Sugar Creek model
bs_confint <- confint(bs_bootstrap, 
                      level = 0.95)

# Conduct bootstrapping on parameter estimates for Elk River model
set.seed(57458) # Set seed so that results are reproducible
elk_bootstrap <- Boot(elk, 
                     R = 9999)

# Save Big Sugar Creek bootstrap data for downstream analysis
save(bs_bootstrap, file = "data/river_data/bs_bootstrap.Rda")

# Save Elk River bootstrap data for downstream analysis
save(elk_bootstrap, file = "data/river_data/elk_bootstrap.Rda")
```

<b>Confidence intervals for Big Sugar Creek parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 417.9443640 - 556.1500369 <br>
<b>K</b>: 0.1771707 - 0.3226116 <br>
<b>t<sub>0</sub></b>: -0.6254293 - -0.1084089 <br>

<b>Confidence intervals for Elk River parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 391.2426135 - 490.53084606 <br>
<b>K</b>: 0.2246298 - 0.39204444 <br>
<b>t<sub>0</sub></b>: -0.5779102 - -0.07034693 <br>

Results from bootstrapping and confidence intervals are given in Table S3b in the final manuscript.

##### 2b.5. Plot histograms of bootstrap results for rivers; run the Rmd chunk below.
In this step, we plot histograms of the bootstrap replicate results to visualize 95% confidence intervals.

##### Plot histograms of bootstrap results for Big Sugar Creek and Elk River: `figures/bootstrap_plots/river.pdf`
```{r}
# Load Big Sugar Creek bootstrap data
load("data/river_data/bs_bootstrap.Rda")

# Load Elk River bootstrap data
load("data/river_data/elk_bootstrap.Rda")

# Convert bootstrap data for Big Sugar Creek to dataframe
bs_boot_data <- data.frame(bs_bootstrap$t)

# Convert bootstrap data for Elk River to dataframe
elk_boot_data <- data.frame(elk_bootstrap$t)

# Create column for Big Sugar Creek identifier in Big Sugar Creek bootstrap dataframe
bs_boot_data$river <- factor(rep("Big_Sugar", rep = 9999))

# Create column for Elk River identifier in Elk River bootstrap dataframe
elk_boot_data$river <- factor(rep("Elk_River", rep = 9999))

# rbind Big Sugar Creek and Elk River bootstrap data
river_boot_data <- rbind(bs_boot_data,
                         elk_boot_data)

# Plot histogram of Linf bootstrap values
Linf <- ggplot() + 
  geom_histogram(data = river_boot_data, aes(x = Linf, fill = river), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 391.2426135, xend = 490.53084606), y = 2100, yend = 2100, color = "grey80", stat = "identity", inherit.aes = F) +
  geom_segment(aes(x = 417.9443640, xend = 556.15003690), y = 1600, yend = 1600, color = "grey20", stat = "identity", inherit.aes = F) +
  geom_point(aes(x = 430.14416, y = 2100), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F, inherit.aes = F) +
  geom_point(aes(x = 464.18395, y = 1600), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F, inherit.aes = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "L∞", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(L[infinity])))

# Plot histogram of K bootstrap values
K <- ggplot() + 
  geom_histogram(data = river_boot_data, aes(x = K, fill = river), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 0.2246298, xend = 0.39204444), y = 1200, yend = 1200, color = "grey80", stat = "identity") +
  geom_segment(aes(x = 0.1771707, xend = 0.32261160), y = 1400,  yend = 1400, color = "grey20", stat = "identity") +
  geom_point(aes(x = 0.30578, y = 1200), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 0.25180, y = 1400), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "K", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.title.x = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(K[B])))

# Plot histogram of t<sub>0</sub> bootstrap values
t0 <- ggplot() +
  geom_histogram(data = river_boot_data, aes(x = t0, fill = river), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = -0.5779102, xend = -0.07034693), y = 1200, yend = 1200, color = "grey80", stat = "identity") +
  geom_segment(aes(x = -0.6254293, xend = -0.10840890), y = 1400, yend = 1400, color = "grey20", stat = "identity") +
  geom_point(aes(x = -0.30943, y = 1200), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = -0.28753, y = 1400), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "t0", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic("t"[0]))) 

# Plot histograms of each parameter estimate together for the river data
pdf("figures/bootstrap_plots/river.pdf", width = 15, height = 4)

plot_grid(Linf, 
          K, 
          t0, 
          nrow = 1, 
          ncol = 3)

dev.off()
```

##### 2b.6. Perform bootstrapping of predicted average length-at-age values and calculate 95% confidence intervals; run the Rmd chunk below.
In this step, we perform bootstrapping over average length-at-age estimates predicted for the Big Sugar Creek and Elk River datasets and calculate 95% confidence intervals around the parameter estimates.

We use the function and sequence of annulus values ("annuli") defined in Step 2a.4 (above). 

##### Conduct bootstrapping over predicted values for rivers:
```{r}
## Conduct bootstrapping and calculated confidence intervals for Big Sugar Creek

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
bs_iboot <- Boot(bs, 
                 f = predict_function,
                 R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
bs_iconfint <- confint(bs_iboot, 
                       level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
bs_predict <- data.frame(annuli,
                         predict(bs, data.frame(annulus = annuli)),
                         bs_iconfint)

# Generate column to distinguish these data as Big Sugar Creek
bs_predict$river <- c(rep("Big_Sugar", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(bs_predict) <- c("annulus","fit","lci","uci", "river")


## Conduct bootstrapping and calculated confidence intervals for Elk River

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
elk_iboot <- Boot(elk, 
                  f = predict_function,
                  R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
elk_iconfint <- confint(elk_iboot, 
                        level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
elk_predict <- data.frame(annuli,
                          predict(elk, data.frame(annulus = annuli)),
                          elk_iconfint)

# Generate column to distinguish these data as Elk River
elk_predict$river <- c(rep("Elk_River", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(elk_predict) <- c("annulus","fit","lci","uci", "river")

## Save predicted datasets for downstream analyses 

# Save the predicted Big Sugar Creek model dataset for donwstream analyses
save(bs_predict, file = "data/river_data/bs_predict.Rda")

# Save the predicted Elk River model dataset for donwstream analyses
save(elk_predict, file = "data/river_data/elk_predict.Rda")
```

##### 2b.7. Plot von Bertlanffy growth for rivers; run the Rmd chunk below.

##### Plot the von Bertalanffy model for rivers: `figures/vb_plots/river.pdf`
```{r}
# Load in Big Sugar Creek predicted dataset to plot best fit curve
load("data/river_data/bs_predict.Rda")

# Load in Elk River predicted dataset to plot best fit curve
load("data/river_data/elk_predict.Rda")

# Load in full back-calculuated data for Big Sugar Creek
load("data/river_data/full_bc_data_bs.Rda")

# Load in full back-calculuated data for Elk River
load("data/river_data/full_bc_data_elk.Rda")

# Plot von Bertalanffy model
pdf("figures/river.pdf", width = 5, height = 4)

ggplot() + 
  geom_ribbon(data = bs_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = river, alpha = 0.2), show.legend = F) +
  geom_ribbon(data = elk_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = river, alpha = 0.2), show.legend = F) +
  geom_hline(yintercept = 464.18395, linetype = "solid", size = 0.5) +
  geom_hline(yintercept = 430.14416, linetype = "dashed", size = 0.5) +
  geom_point(data = full_bc_data_bs, aes(x = annulus, y = bc_tl, fill = river, group = river, shape = river), show.legend = F, size = 3, pch = 22, position = position_jitter(width = 0.4)) + 
  geom_point(data = full_bc_data_elk, aes(x = annulus, y = bc_tl, fill = river, group = river, shape = river), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.2)) + 
  geom_line(data = bs_predict, aes(x = annulus, y = fit, linetype = river), show.legend = F) +
  geom_line(data = elk_predict, aes(x = annulus, y = fit, linetype = river), show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("grey20","grey80")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,550), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

This figure is the basis for Figure 2b in the final manuscript.

#### 2c: Generate growth model for each sex (Male, Female) separately.
In this step, we parameterize the von Bertalanffy growth model for each sex (Male, Female) to assess potential differences in growth due to sexual dimorphism.

##### 2c.1. Load full back-calculation data and separate into datasets by sex; run the Rmd chunk below.

##### Separate global dataset by sex:
```{r}
# Load full bc data
load("data/bc_data/full_bc_data.Rda")

# Get male samples only
full_bc_data_m <- full_bc_data %>%
  filter(sex == "Male")

# Get female samples only
full_bc_data_f <- full_bc_data %>%
  filter(sex == "Female")

# save male data for downstream analyses
save(full_bc_data_m, file = "data/sex_data/full_bc_data_m.Rda")

# save female data for downstream analyses
save(full_bc_data_f, file = "data/sex_data/full_bc_data_f.Rda")
```

##### 2c.2. Generate model starting values for each sex; run the Rmd chunk below.

##### load data and generate starting values for sex:
```{r}
# Load full bc data for males
load("data/sex_data/full_bc_data_m.Rda")

# Load full bc data for females
load("data/sex_data/full_bc_data_f.Rda")

# Generate starting values for optimizing the von Bertalanffy model in males
m_starts <- vbStarts(bc_tl ~ annulus, 
                     data = full_bc_data_m)

# Generate starting values for optimizing the von Bertalanffy model in females
f_starts <- vbStarts(bc_tl ~ annulus, 
                     data = full_bc_data_f)
```

<b>Starting values used for male model</b>: <br>

<b>L<sub>inf</sub></b>: 436.1392 <br>
<b>K</b>: 0.3052123 <br>
<b>t<sub>0</sub></b>: -0.2546104 <br>

<b>Starting values used for female model</b>: <br>

<b>L<sub>inf</sub></b>: 383.0877 <br>
<b>K</b>: 0.3944219 <br>
<b>t<sub>0</sub></b>: -0.09356921 <br>

##### 2c.3. Model growth for each sex and extract model coefficients; run the Rmd chunk below.
In this step, we run the non-linear von Bertalanffy model for each sex and extract model coefficients from the summary table given with the summary() function.

##### Model von Bertalanffy growth for sex:
```{r}
# Run von Bertalanffy growth model for males using back-calculated total length, modeled by annulus number (consensus age)
m <- nls(bc_tl ~ vb(annulus,
                    Linf,
                    K,
                    t0), 
         data = full_bc_data_m, 
         start = m_starts)

# Extract male data coefficients
m_coefficients <- summary(m)

# Run von Bertalanffy growth model for females using back-calculated total length, modeled by annulus number (consensus age)
f <- nls(bc_tl ~ vb(annulus,
                    Linf,
                    K,
                    t0), 
         data = full_bc_data_f, 
         start = f_starts)

# Extract female data coefficients
f_coefficients <- summary(f)
```

<b>Parameter estimates for male model</b>: <br>

<b>L<sub>inf</sub></b>: 445.5611 <br>
<b>K</b>: 0.2894 <br>
<b>t<sub>0</sub></b>: -0.2842 <br>

<b>Parameter estimates for female model</b>: <br>

<b>L<sub>inf</sub></b>: 440.35229 <br>
<b>K</b>: 0.27610 <br>
<b>t<sub>0</sub></b>: -0.31403 <br>

Parameter estimates are given in Table S3c in the final manuscript.

##### 2c.4. Perform bootstrapping on sex datasets to estimate 95% confidence intervals around each parameter estimate; run the Rmd chunk below.
In this step, we perform bootstrapping on each sex with 9999 bootstrap replicates and calculate 95% confidence intervals around the parameter estimates given in the male and female growth models (see Step 2c.3).

##### Perform bootstrapping on male and female models:
```{r}
# Conduct bootstrapping on parameter estimates for male model
set.seed(53744) # Set seed so that results are reproducible
m_bootstrap <- Boot(m, 
                    R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for male model
m_confint <- confint(m_bootstrap, 
                     level = 0.95)

# Conduct bootstrapping on parameter estimates for female model
set.seed(340985) # Set seed so that results are reproducible
f_bootstrap <- Boot(f, 
                    R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for female model
f_confint <- confint(f_bootstrap, 
                     level = 0.95)

# Save male bootstrap data for downstream analysis
save(m_bootstrap, file = "data/sex_data/m_bootstrap.Rda")

# Save female bootstrap data for downstream analysis
save(f_bootstrap, file = "data/sex_data/f_bootstrap.Rda")
```

<b>Confidence intervals for male parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 407.1501469 - 511.6867181 <br>
<b>K</b>: 0.2137145 - 0.3718532 <br>
<b>t<sub>0</sub></b>: -0.5874782 - -0.0678828 <br>

<b>Confidence intervals for female parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 396.3111777 - 515.08685844 <br>
<b>K</b>: 0.1967684 - 0.35531175 <br>
<b>t<sub>0</sub></b>: -0.6221602 - -0.09637011 <br>

Results from bootstrapping and confidence intervals are given in Table S3c in the final manuscript.

##### 2c.5. Plot histograms of bootstrap results for sexes; run the Rmd chunk below.
In this step, we plot histograms of the bootstrap replicate results to visualize 95% confidence intervals.

##### Plot histograms of bootstrap results for males and females: `figures/bootstrap_plots/sex.pdf`
```{r}
# Load male bootstrap data
load("data/sex_data/m_bootstrap.Rda")

# Load female bootstrap data
load("data/sex_data/f_bootstrap.Rda")

# Convert bootstrap data for males to dataframe
m_boot_data <- data.frame(m_bootstrap$t)

# Convert bootstrap data for females to dataframe
f_boot_data <- data.frame(f_bootstrap$t)

# Create column for female identifier in female bootstrap dataframe
m_boot_data$sex <- factor(rep("Male", rep = 9999))

# Create column for female identifier in male bootstrap dataframe
f_boot_data$sex <- factor(rep("Female", rep = 9999))

# rbind male and female bootstrap data
sex_boot_data <- rbind(m_boot_data,
                       f_boot_data)

# Reorder levels so that histogram colors plot in the right order (light gray on top)
sex_boot_data$sex = factor(sex_boot_data$sex, levels = c("Female","Male"))

# Plot histogram of Linf bootstrap values
Linf <- ggplot() + 
  geom_histogram(data = sex_boot_data, aes(x = Linf, fill = sex), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 407.1501469, xend = 511.68671810), y = 2300, yend = 2300, color = "grey80", stat = "identity") +
  geom_segment(aes(x = 396.3111777, xend = 515.08685844), y = 2050, yend = 2050, color = "grey20", stat = "identity") +
  geom_point(aes(x = 445.56110, y = 2300), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 440.35229, y = 2050), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "L∞", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(L[infinity])))

# Plot histogram of K bootstrap values
K <- ggplot() + 
  geom_histogram(data = sex_boot_data, aes(x = K, fill = sex), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 0.2137145, xend = 0.37185320), y = 1200, yend = 1200, color = "grey80", stat = "identity") +
  geom_segment(aes(x = 0.1967684, xend = 0.35531175), y = 1100,  yend = 1100, color = "grey20", stat = "identity") +
  geom_point(aes(x = 0.28940, y = 1200), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 0.27610, y = 1100), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "K", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.title.x = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(K[B])))

# Plot histogram of t<sub>0</sub> bootstrap values
t0 <- ggplot() + 
  geom_histogram(data = sex_boot_data, aes(x = t0, fill = sex), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = -0.5874782, xend = -0.06788280), y = 1300, yend = 1300, color = "grey80", stat = "identity") +
  geom_segment(aes(x = -0.6221602, xend = -0.09637011), y = 1200, yend = 1200, color = "grey20", stat = "identity") +
  geom_point(aes(x = -0.28420, y = 1300), fill = "grey80", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = -0.31403, y = 1200), fill = "grey20", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "t0", y = "Frequency") +
  scale_fill_manual(values = c("grey20","grey80")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic("t"[0])))

# Plot histograms of each parameter estimate together for the sex data
pdf("figures/bootstrap_plots/sex.pdf", width = 15, height = 4)

plot_grid(Linf, 
          K, 
          t0, 
          nrow = 1, 
          ncol = 3)

dev.off()
```

##### 2c.6. Perform bootstrapping of predicted average length-at-age values and calculate 95% confidence intervals; run the Rmd chunk below.
In this step, we perform bootstrapping over average length-at-age estimates predicted for the male and female datasets and calculate 95% confidence intervals around the parameter estimates.

We use the function and sequence of annulus values ("annuli") defined in Step 2a.4 (above). 

##### Conduct bootstrapping over predicted values:
```{r}
## Conduct bootstrapping and calculated confidence intervals for males

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
m_iboot <- Boot(m, 
                f = predict_function,
                R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
m_iconfint <- confint(m_iboot, 
                      level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
m_predict <- data.frame(annuli,
                        predict(m, data.frame(annulus = annuli)),
                        m_iconfint)

# Generate column to distinguish these data as male
m_predict$sex <- c(rep("Male", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(m_predict) <- c("annulus","fit","lci","uci", "sex")


## Conduct bootstrapping and calculated confidence intervals for females

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
f_iboot <- Boot(f, 
                f = predict_function,
                R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
f_iconfint <- confint(f_iboot, 
                      level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
f_predict <- data.frame(annuli,
                        predict(f, data.frame(annulus = annuli)),
                        f_iconfint)

# Generate column to distinguish these data as female
f_predict$sex <- c(rep("Female", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(f_predict) <- c("annulus","fit","lci","uci", "sex")

## Save predicted datasets for downstream analyses 

# Save the predicted male model dataset for donwstream analyses
save(m_predict, file = "data/sex_data/m_predict.Rda")

# Save the predicted female model dataset for donwstream analyses
save(f_predict, file = "data/sex_data/f_predict.Rda")
```

##### 2c.7. Plot von Bertlanffy growth for sex; run the Rmd chunk below.

##### Plot the von Bertalanffy model for sex: `figures/vb_plots/sex.pdf`
```{r}
# Load in male predicted dataset to plot best fit curve
load("data/sex_data/m_predict.Rda")

# Load in female predicted dataset to plot best fit curve
load("data/sex_data/f_predict.Rda")

# Load in full back-calculated data for males
load("data/sex_data/full_bc_data_m.Rda")

# Load in full back-calculated data for females
load("data/sex_data/full_bc_data_f.Rda")

# Plot von Bertalanffy model
pdf("figures/sex.pdf", width = 5, height = 4)

ggplot() + 
  geom_ribbon(data = m_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = sex, alpha = 0.2), show.legend = F) +
  geom_ribbon(data = f_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = sex, alpha = 0.2), show.legend = F) +
  geom_hline(yintercept = 445.5611, linetype = "dashed", size = 0.5) +
  geom_hline(yintercept = 440.35229, linetype = "solid", size = 0.5) +
  geom_point(data = full_bc_data_m, aes(x = annulus, y = bc_tl, fill = sex, group = sex, shape = sex), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.4)) + 
  geom_point(data = full_bc_data_f, aes(x = annulus, y = bc_tl, fill = sex, group = sex, shape = sex), show.legend = F, size = 3, pch = 22, position = position_jitter(width = 0.2)) + 
  geom_line(data = m_predict, aes(x = annulus, y = fit, linetype = sex), show.legend = F) +
  geom_line(data = f_predict, aes(x = annulus, y = fit, linetype = sex), show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("grey20","grey80")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,550), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size=1)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

This figure is the basis for Figure 2c in the final manuscript.

#### 2d: Generate growth model for each ancestry group (Smallmouth Bass (SMB), Neosho Bass (NB), and Admixed (ADM)) separately.
In this step, we parameterize the von Bertalanffy growth model for each ancestry group (Smallmouth Bass (SMB), Neosho Bass (NB), and Admixed (ADM)) to assess potential differences in growth due to pure or mixed ancestry.

##### 2d.1. Load full back-calculation data and separate into datasets by ancestry group; run the Rmd chunk below.

##### Separate global dataset by sex:
```{r}
# Load full bc data
load("data/bc_data/full_bc_data.Rda")

# Get NB samples only
full_bc_data_nb <- full_bc_data %>%
  filter(ancestry_group == "Neosho_Bass")

# Get SMB samples only
full_bc_data_smb <- full_bc_data %>%
  filter(ancestry_group == "Smallmouth_Bass")

# Get ADM samples only
full_bc_data_adm <- full_bc_data %>%
  filter(ancestry_group == "Admixed")

# save NB data for downstream analyses
save(full_bc_data_nb, file = "data/ancestry_data/full_bc_data_nb.Rda")

# save SMB data for downstream analyses
save(full_bc_data_smb, file = "data/ancestry_data/full_bc_data_smb.Rda")

# save ADM data for downstream analyses
save(full_bc_data_adm, file = "data/ancestry_data/full_bc_data_adm.Rda")
```

##### 2d.2. Generate model starting values for each ancestry group; run the Rmd chunk below.

##### Load data and generate starting values for ancestry group:
```{r}
# Load full bc data for NB
load("data/ancestry_data/full_bc_data_nb.Rda")

# Load full bc data for SMB
load("data/ancestry_data/full_bc_data_smb.Rda")

# Load full bc data for ADM
load("data/ancestry_data/full_bc_data_adm.Rda")

# Generate starting values for optimizing the von Bertalanffy model in NB
nb_starts <- vbStarts(bc_tl ~ annulus, 
                      data = full_bc_data_nb)

# Generate starting values for optimizing the von Bertalanffy model in SMB
smb_starts <- vbStarts(bc_tl ~ annulus, 
                       data = full_bc_data_smb)

# Generate starting values for optimizing the von Bertalanffy model in ADM
adm_starts <- vbStarts(bc_tl ~ annulus, 
                       data = full_bc_data_adm)
```

<b>Starting values used for NB model</b>: <br>

<b>L<sub>inf</sub></b>: 445.85 <br>
<b>K</b>: 0.2884396 <br>
<b>t<sub>0</sub></b>: -0.2720177 <br>

<b>Starting values used for SMB model</b>: <br>

<b>L<sub>inf</sub></b>: 526.263 <br>
<b>K</b>: 0.2119943 <br>
<b>t<sub>0</sub></b>: -0.2800102 <br>

<b>Starting values used for ADM model</b>: <br>

<b>L<sub>inf</sub></b>: 375.3536 <br>
<b>K</b>: 0.4146585 <br>
<b>t<sub>0</sub></b>: -0.1198689 <br>

##### 2d.3. Model growth for each ancestry group and extract model coefficients; run the Rmd chunk below.
In this step, we run the non-linear von Bertalanffy model for each ancestry and extract model coefficients from the summary table given with the summary() function.

##### Model von Bertalanffy growth for sex:
```{r}
# Run von Bertalanffy growth model for males using back-calculated total length, modeled by annulus number (consensus age)
nb <- nls(bc_tl ~ vb(annulus,
                     Linf,
                     K,
                     t0), 
         data = full_bc_data_nb, 
         start = nb_starts)

# Extract male data coefficients
nb_coefficients <- summary(nb)

# Run von Bertalanffy growth model for SMB using back-calculated total length, modeled by annulus number (consensus age)
smb <- nls(bc_tl ~ vb(annulus,
                      Linf,
                      K,
                      t0), 
         data = full_bc_data_smb, 
         start = smb_starts)

# Extract female data coefficients
smb_coefficients <- summary(smb)

# Run von Bertalanffy growth model for ADM using back-calculated total length, modeled by annulus number (consensus age)
adm <- nls(bc_tl ~ vb(annulus,
                      Linf,
                      K,
                      t0), 
         data = full_bc_data_adm, 
         start = adm_starts)

# Extract female data coefficients
adm_coefficients <- summary(adm)
```

<b>Parameter estimates for NB model</b>: <br>

<b>L<sub>inf</sub></b>: 487.23225 <br>
<b>K</b>: 0.23643 <br>
<b>t<sub>0</sub></b>: -0.39424 <br>

<b>Parameter estimates for SMB model</b>: <br>

<b>L<sub>inf</sub></b>: 542.33111 <br>
<b>K</b>: 0.19377 <br>
<b>t<sub>0</sub></b>: -0.37020 <br>

<b>Parameter estimates for ADM model</b>: <br>

<b>L<sub>inf</sub></b>: 403.55265 <br>
<b>K</b>: 0.33793 <br>
<b>t<sub>0</sub></b>: -0.24880 <br>

Parameter estimates are given in Table S3d in the final manuscript.

##### 2d.4. Perform bootstrapping on ancestry group datasets to estimate 95% confidence intervals around each parameter estimate; run the Rmd chunk below.
In this step, we perform bootstrapping on each sex with 9999 bootstrap replicates and calculate 95% confidence intervals around the parameter estimates given in the ancestry group growth models (see Step 2d.3).

##### Perform bootstrapping on ancestry group models:
```{r}
# Conduct bootstrapping on parameter estimates for NB model
set.seed(47534) # Set seed so that results are reproducible
nb_bootstrap <- Boot(nb, 
                     R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for NB model
nb_confint <- confint(nb_bootstrap, 
                     level = 0.95)

# Conduct bootstrapping on parameter estimates for SMB model
set.seed(39084) # Set seed so that results are reproducible
smb_bootstrap <- Boot(smb, 
                      R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for ADM model
smb_confint <- confint(smb_bootstrap, 
                       level = 0.95)

# Conduct bootstrapping on parameter estimates for ADM model
set.seed(37822) # Set seed so that results are reproducible
adm_bootstrap <- Boot(adm, 
                      R = 9999)

# Calculate bootstrapped 95% confidence intervals for parameter estimates for ADM model
adm_confint <- confint(adm_bootstrap, 
                       level = 0.95)

# Save NB bootstrap data for downstream analysis
save(nb_bootstrap, file = "data/ancestry_data/nb_bootstrap.Rda")

# Save SMB bootstrap data for downstream analysis
save(smb_bootstrap, file = "data/ancestry_data/smb_bootstrap.Rda")

# Save ADM bootstrap data for downstream analysis
save(adm_bootstrap, file = "data/ancestry_data/adm_bootstrap.Rda")
```

<b>Confidence intervals for NB parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 431.9843230 - 597.3864203 <br>
<b>K</b>: 0.1581719 - 0.3204952 <br>
<b>t<sub>0</sub></b>: -0.7711496 - -0.1205203 <br>

<b>Confidence intervals for SMB parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 323.75679667 - 1259.6571567 <br>
<b>K</b>: 0.05444464 - 0.5235279 <br>
<b>t<sub>0</sub></b>: -1.10145867 - -0.1077478 <br>

<b>Confidence intervals for ADM parameter estimates</b>: <br>

<b>L<sub>inf</sub></b>: 371.6570318 - 447.92298437 <br>
<b>K</b>: 0.2573298 - 0.42598624 <br>
<b>t<sub>0</sub></b>: -0.5342975 - -0.04283142 <br>

Results from bootstrapping and confidence intervals are given in Table S3D in the final manuscript.

##### 2d.5. Plot histograms of bootstrap results for ancestry; run the Rmd chunk below.
In this step, we plot histograms of the bootstrap replicate results to visualize 95% confidence intervals.

##### Plot histograms of bootstrap results for ancestry groups: `figures/bootstrap_plots/ancestry.pdf`
```{r}
# Load NB bootstrap data
load("data/ancestry_data/nb_bootstrap.Rda")

# Load SMB bootstrap data
load("data/ancestry_data/smb_bootstrap.Rda")

# Load ADM bootstrap data
load("data/ancestry_data/adm_bootstrap.Rda")

# Convert bootstrap data for NB to dataframe
nb_boot_data <- data.frame(nb_bootstrap$t)

# Convert bootstrap data for SMB to dataframe
smb_boot_data <- data.frame(smb_bootstrap$t)

# Convert bootstrap data for ADM to dataframe
adm_boot_data <- data.frame(adm_bootstrap$t)

# Create column for NB identifier in NB bootstrap dataframe
nb_boot_data$ancestry <- factor(rep("NB", rep = 9999))

# Create column for SMB identifier in SMB bootstrap dataframe
smb_boot_data$ancestry <- factor(rep("SMB", rep = 9999))

# Create column for ADM identifier in ADM bootstrap dataframe
adm_boot_data$ancestry <- factor(rep("ADM", rep = 9999))

# rbind ancestry group bootstrap data
ancestry_boot_data <- rbind(nb_boot_data,
                            smb_boot_data,
                            adm_boot_data)

# Plot histogram of Linf bootstrap values
Linf <- ggplot() + 
  geom_histogram(data = ancestry_boot_data, aes(x = Linf, fill = ancestry), color = "black", alpha = 0.8, position = "identity", show.legend = F, bins = 40) +
  geom_segment(aes(x = 431.98432300, xend = 597.38642030), y = 2200, yend = 2200, color = "deepskyblue", stat = "identity") +
  geom_segment(aes(x = 371.65703180, xend = 447.92298437), y = 4200, yend = 4200, color = "mediumpurple", stat = "identity") +
  geom_segment(aes(x = 323.75679667, xend = 1259.6571567), y = 800, yend = 800, color = "deeppink2", stat = "identity") +
  geom_point(aes(x = 487.23225, y = 2200), fill = "deepskyblue", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 403.55265, y = 4200), fill = "mediumpurple", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 542.33111, y = 800), fill = "deeppink2", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "L∞", y = "Frequency") +
  scale_fill_manual(values = c("deepskyblue","deeppink2","mediumpurple")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(L[infinity]))) +
  xlim(150,1260)

# Plot histogram of K bootstrap values
K <- ggplot() + 
  geom_histogram(data = ancestry_boot_data, aes(x = K, fill = ancestry), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = 0.15817190, xend = 0.32049520), y = 1600, yend = 1600, color = "deepskyblue", stat = "identity") +
  geom_segment(aes(x = 0.25732980, xend = 0.42598624), y = 1500, yend = 1500, color = "mediumpurple", stat = "identity") +
  geom_segment(aes(x = 0.05444464, xend = 0.52352790), y = 700, yend = 700, color = "deeppink2", stat = "identity") +
  geom_point(aes(x = 0.23643, y = 1600), fill = "deepskyblue", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 0.33793, y = 1500), fill = "mediumpurple", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = 0.19377, y = 700), fill = "deeppink2", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "K", y = "Frequency") +
  scale_fill_manual(values = c("deepskyblue","deeppink2","mediumpurple")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.title.x = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic(K[B]))) +
  xlim(0,0.53)

# Plot histogram of t<sub>0</sub> bootstrap values
t0 <- ggplot() + 
  geom_histogram(data = ancestry_boot_data, aes(x = t0, fill = ancestry), color = "black", alpha = 0.8, position = "identity", show.legend = F) +
  geom_segment(aes(x = -0.77114960, xend = -0.12052030), y = 1500, yend = 1500, color = "deepskyblue", stat = "identity") +
  geom_segment(aes(x = -0.53429750, xend = -0.04283142), y = 1900, yend = 1900, color = "mediumpurple", stat = "identity") +
  geom_segment(aes(x = -1.10145867, xend = -0.10774780), y = 800, yend = 800, color = "deeppink2", stat = "identity") +
  geom_point(aes(x = -0.39424, y = 1500), fill = "deepskyblue", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = -0.24880, y = 1900), fill = "mediumpurple", color = "black", pch = 21, size = 4, show.legend = F) +
  geom_point(aes(x = -0.37020, y = 800), fill = "deeppink2", color = "black", pch = 21, size = 4, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  labs(x = "t0", y = "Frequency") +
  scale_fill_manual(values = c("deepskyblue","deeppink2","mediumpurple")) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme(plot.margin = margin(10,10,10,10)) +
  xlab(expression(italic("t"[0]))) +
  xlim(-1.5,0.5)

# Plot histograms of each parameter estimate together for the sex data
pdf("figures/bootstrap_plots/ancestry.pdf", width = 15, height = 4)

plot_grid(Linf, 
          K, 
          t0, 
          nrow = 1, 
          ncol = 3)

dev.off()
```

##### 2d.6. Perform bootstrapping of predicted average length-at-age values and calculate 95% confidence intervals; run the Rmd chunk below.
In this step, we perform bootstrapping over average length-at-age estimates predicted for the ancestry group datasets and calculate 95% confidence intervals around the parameter estimates.

We use the function and sequence of annulus values ("annuli") defined in Step 2a.4 (above). 

##### Conduct bootstrapping over predicted values for ancestry group:
```{r}
## Conduct bootstrapping and calculated confidence intervals for NB

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
nb_iboot <- Boot(nb, 
                 f = predict_function,
                 R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
nb_iconfint <- confint(nb_iboot, 
                       level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
nb_predict <- data.frame(annuli,
                         predict(nb, data.frame(annulus = annuli)),
                         nb_iconfint)

# Generate column to distinguish these data as male
nb_predict$ancestry_group <- c(rep("Neosho_Bass", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(nb_predict) <- c("annulus","fit","lci","uci", "ancestry_group")


## Conduct bootstrapping and calculated confidence intervals for SMB

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
smb_iboot <- Boot(smb, 
                  f = predict_function,
                  R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
smb_iconfint <- confint(smb_iboot, 
                        level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
smb_predict <- data.frame(annuli,
                          predict(smb, data.frame(annulus = annuli)),
                          smb_iconfint)

# Generate column to distinguish these data as female
smb_predict$ancestry_group <- c(rep("Smallmouth_Bass", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(smb_predict) <- c("annulus","fit","lci","uci", "ancestry_group")


## Conduct bootstrapping and calculated confidence intervals for ADM

# Conduct bootstrapping with 9999 bootstrap replicates to calculate confidence intervals around the best fit curve
adm_iboot <- Boot(adm, 
                  f = predict_function,
                  R = 9999)

# Calculate bootstrapped 95% confidence intervals for all parameter estimates
adm_iconfint <- confint(adm_iboot, 
                        level = 0.95)

# Predict average back-calculated length-at-age for all fish based on model fit and confidence intervals
adm_predict <- data.frame(annuli,
                          predict(adm, data.frame(annulus = annuli)),
                          adm_iconfint)

# Generate column to distinguish these data as female
adm_predict$ancestry_group <- c(rep("Admixed", times = 81))

# Modify column names (lci = lower confidence interval [2.5%]; uci = upper confidence interval [97.5%])
colnames(adm_predict) <- c("annulus","fit","lci","uci", "ancestry_group")


## Save predicted datasets for downstream analyses 

# Save the predicted NB model dataset for donwstream analyses
save(nb_predict, file = "data/ancestry_data/nb_predict.Rda")

# Save the predicted SMB model dataset for donwstream analyses
save(smb_predict, file = "data/ancestry_data/smb_predict.Rda")

# Save the predicted ADM model dataset for donwstream analyses
save(adm_predict, file = "data/ancestry_data/adm_predict.Rda")
```

##### 2d.7. Plot von Bertlanffy growth for ancestry group; run the Rmd chunk below.

##### Plot the von Bertalanffy model for ancestry group: `figures/vb_plots/ancestry.pdf`
```{r}
# Load in NB predicted dataset to plot best fit curve
load("data/ancestry_data/nb_predict.Rda")

# Load in SMB predicted dataset to plot best fit curve
load("data/ancestry_data/smb_predict.Rda")

# Load in ADM predicted dataset to plot best fit curve
load("data/ancestry_data/adm_predict.Rda")

# Load in full back-calculated data for NB
load("data/ancestry_data/full_bc_data_nb.Rda")

# Load in full back-calculated data for SMB
load("data/ancestry_data/full_bc_data_smb.Rda")

# Load in full back-calculated data for ADM
load("data/ancestry_data/full_bc_data_adm.Rda")

# Plot von Bertalanffy model
pdf("figures/vb_plots/ancestry.pdf", width = 6, height = 4)

ggplot() + 
  geom_hline(yintercept = 487.23225, linetype = "dotted", size = 0.8, color = "deepskyblue") +
  geom_hline(yintercept = 403.55265, linetype = "solid", size = 0.8, color = "mediumpurple") +
  geom_hline(yintercept = 542.33111, linetype = "longdash", size = 0.8, color = "deeppink2") +
  geom_point(data = full_bc_data_nb, aes(x = annulus, y = bc_tl, fill = ancestry_group, group = ancestry_group, shape = ancestry_group), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.4)) + 
  geom_point(data = full_bc_data_smb, aes(x = annulus, y = bc_tl, fill = ancestry_group, group = ancestry_group, shape = ancestry_group), show.legend = F, size = 3, pch = 22, position = position_jitter(width = 0.2)) + 
  geom_point(data = full_bc_data_adm, aes(x = annulus, y = bc_tl, fill = ancestry_group, group = ancestry_group, shape = ancestry_group), show.legend = F, size = 3, pch = 24, position = position_jitter(width = 0.4)) + 
  geom_line(data = nb_predict, aes(x = annulus, y = fit, linetype = ancestry_group, color = ancestry_group), show.legend = F) +
  geom_line(data = smb_predict, aes(x = annulus, y = fit, linetype = ancestry_group, color = ancestry_group), show.legend = F) +
  geom_line(data = adm_predict, aes(x = annulus, y = fit, linetype = ancestry_group, color = ancestry_group), show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("mediumpurple","deepskyblue","deeppink2")) +
  scale_color_manual(values = c("mediumpurple","deepskyblue","deeppink2")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,575), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size=1)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

We initially plotted all three ancestry group growth models, including best lines, confidence interval ribbons around the best fit lines, and observed data points, on the same plot. The confidence intervals for Smallmouth Bass (deeppink) were substantially greater than those for Neosho Bass (deepskyblue) and Admixed individuls (mediumpurple) and obscured view of the Neosho Bass and Admixed growth curves. For this reason, we omitted the confidence intervals for this plot and included only the real data points and best fit lines. 

This figure is the basis for Figure S[...] in the final manuscript.

##### 2d.8. Plot von Bertlanffy growth for Neosho Bass (NB) and Admixed (ADM) ancestry groups; run the Rmd chunk below.
In this step, we plot the same data as above (Step 2d.6), but we include only the data points and associated best fit curves and confidence interval ribbons for the NB and ADM curves for better viewing. We omit data for Smallmouth Bass from the main curve, and instead show the confidence intervals for SMB in an inset plot.

##### Plot the von Bertalanffy model for ancestry group: `figures/vb_plots/nb_adm_ancestry.pdf`
```{r}
# Load in NB predicted dataset to plot best fit curve
load("data/ancestry_data/nb_predict.Rda")

# Load in ADM predicted dataset to plot best fit curve
load("data/ancestry_data/adm_predict.Rda")

# Load in full back-calculated data for NB
load("data/ancestry_data/full_bc_data_nb.Rda")

# Load in full back-calculated data for ADM
load("data/ancestry_data/full_bc_data_adm.Rda")

# Plot von Bertalanffy model
pdf("figures/nb_adm_ancestry.pdf", width = 5, height = 4)

ggplot() + 
  geom_ribbon(data = nb_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = ancestry_group, alpha = 0.2), show.legend = F) +
  geom_ribbon(data = adm_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = ancestry_group, alpha = 0.2), show.legend = F) +
  geom_hline(yintercept = 487.23225, linetype = "dashed", size = 0.5) +
  geom_hline(yintercept = 403.55265, linetype = "solid", size = 0.5) +
  geom_point(data = full_bc_data_nb, aes(x = annulus, y = bc_tl, fill = ancestry_group, group = ancestry_group, shape = ancestry_group), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.4)) + 
  geom_point(data = full_bc_data_adm, aes(x = annulus, y = bc_tl, fill = ancestry_group, group = ancestry_group, shape = ancestry_group), show.legend = F, size = 3, pch = 22, position = position_jitter(width = 0.2)) + 
  geom_line(data = nb_predict, aes(x = annulus, y = fit, linetype = ancestry_group), show.legend = F) +
  geom_line(data = adm_predict, aes(x = annulus, y = fit, linetype = ancestry_group), show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("mediumpurple","deepskyblue")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,575), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size=1)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

This figure is the basis for Figure 2d in the final manuscript.

##### 2d.9. Plot von Bertlanffy growth for SMB ancestry group to be included as an inset in Figure 2d (see Step 2d.7 above); run the Rmd chunk below.
In this step, we plot growth data for Smallmouth Bass only, including real data points, best fit line, and confidence ribbon and include it as in an inset in Figure 2d.

##### Plot the von Bertalanffy model for ancestry group: `figures/vb_plots/smb_ancestry.pdf`
```{r}
# Load in SMB predicted dataset to plot best fit curve
load("data/ancestry_data/smb_predict.Rda")

# Load in full back-calculated data for SMB
load("data/ancestry_data/full_bc_data_smb.Rda")

# Plot von Bertalanffy model
pdf("figures/smb_ancestry.pdf", width = 8, height = 7)

ggplot() + 
  geom_ribbon(data = smb_predict, aes(x = annulus, ymin = lci, ymax = uci, fill = ancestry_group, alpha = 0.2), show.legend = F) +
  geom_hline(yintercept = 542.33111, linetype = "solid", size = 0.8) +
  geom_point(data = full_bc_data_smb, aes(x = annulus, y = bc_tl, fill = ancestry_group), show.legend = F, size = 3, pch = 21, position = position_jitter(width = 0.2)) + 
  geom_line(data = smb_predict, aes(x = annulus, y = fit), show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("deeppink2")) +
  scale_y_continuous(name = "Back-calculated total length (mm)", limits = c(0,800), expand = c(0,0)) +
  scale_x_continuous(name = "Annulus (years)", expand = c(0,0), limits = c(0,15), breaks = seq(0,15,1)) +
  theme(axis.title = element_text(size = 25)) + 
  theme(axis.text = element_text(size = 25)) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size=1)) +
  theme(plot.margin = margin(15,15,15,15))

dev.off()
```

This figure is the basis for Figure S4 in the final manuscript.

### STEP 2: Calculate standardized multi-locus heterozygosity
In this step, I read in the fully filtered raw genotype data and calculate standardized multi-locus heterozygosity, which is individual heterozygosity divided by the global average heterozygosity of the entire dataset. I check for identity disequilibrium among loci by checking for correlation between heterozygous genotypes, which provides support for whether there is sufficient power in the data to detect a heterozygosity-fitness correlation.

#### 2a: Filter the raw data to obtain only genotype data and convert to csv format; run the Rmd chunk below.
In this step, I read in the fully filtered raw data, filter to only the genotype data, and clean for use in the inbreedR package. 

##### Read in full raw data filter genotype data: 
```{r}
# Load in full ancestry data curated in Analysis 3 
load("../ancestry_analysis/data/processed_ancestry_data/full_ancestry_data.Rda")

# Filter out only sample id and genotype columns
genotype_data <- full_ancestry_data %>%
  filter(population == "sample") %>%
  select(1,11:38)

# Convert first column to rownames
genotype_data <- column_to_rownames(genotype_data, "sample_id")

# Convert all instances of missing genotype data ("OOO") to "NA" 
genotype_data[genotype_data == "000"] <- "NA"

# Rename all columns
colnames(genotype_data) <- c("mdo9","NA","mdo5","NA","mdo7","NA","mdo10","NA","mdo6","NA","mdo3","NA","mdo2","NA","lma21","NA","msaf29","NA","msaf05","NA","msaf14","NA","msaf17","NA","msaf09","NA","msaf06","NA")

# Save genotype data for downstream analyses
save(genotype_data, file = "data/genotype_data.Rda")
```

#### 2b: Convert genotype data to inbreedR compatible format; run the Rmd chunk below.
In this step, I read in the formatted file generated in Step 2a above and convert it to the appropriate format for the inbreedR package using the built-in "convert_raw()" function.

##### Convert genotype data to inbreedR format and calculate g2 statistics: 
```{r}
# Load genotype data 
load("data/genotype_data.Rda")

# Convert genotype data to inbreedR format
i_data <- convert_raw(genotype_data)

# Use the built-in "check_data()" function in inbreedR to ensure the data are in the correct format
check_data(i_data) ## data are in correct format

# Save inbreedR data for downstream analyses
save(i_data, file = "data/i_data.Rda")
```


#### 2c: Calculate g2 statistic; run the Rmd chunk below.
In this step, I calculate the g2 identity disequilibrium statistic for the full dataset to determine whether the data have sufficient power for detecting a heterozygosity-fitness correlation.

Parameters for g2 statistic calculation:

<b>Permutations</b>= 1,000 <br>
<b>Bootstrap replicates</b>= 1,000
<b>95% Confidence Intervals</b>

##### Calculate g2 identity disequilibrium statistic for microsatellite loci
```{r}
# Load inbreedR formatted data
load("data/i_data.Rda")

# Calculate g2 statistic for microsatellite data
g2 <- g2_microsats(i_data, 
                   nperm = 1000, 
                   nboot = 1000, 
                   CI = 0.95)
```

<b>Summary of g2 statistical test</b>: <br>

<b><i>g2</i></b>= 0.01014848 <br>
<b><i>se</i></b>= 0.005280294 <br>
<b><i>p(g2 > 0)</i></b>= 0.017 <br>

There is significant power for detecting heterozygosity-fitness correlation if one exists for these data.

#### 2d: Calculate standardized multi-locus heterozygosity; run the Rmd chunk below.
In this step, we use the genotype data formatted in step 2a above to calculate standardized multi-locus heterozygosity for each fish, defined as the individual multi-locus heterozygosity divided by the average heterozygosity of the full dataset.

##### Calculate standardized multi-locus heterozygosity and bind to the condition dataset: 
```{r}
# Calculate standardized multi-locus heterozygosity using the "sMLH()" function
h <- data.frame(sMLH(i_data))

# Modify first column to have a column name
h <- rownames_to_column(h)

# Modify column names for data merging
colnames(h) <- c("sample_id", "h")

## Join h data with full condition data

# Load condition data
load("data/condition_data.Rda")

# Bind condition data with h data for downstream analyses
hfc_data <- merge(condition_data, 
                  h, 
                  by = "sample_id")

# Save hfc data for downstream analyses
save(hfc_data, file = "data/hfc_data.Rda")
```
