---
title: "Analysis 5: Ancestry condition correlation analysis"
author: "Joe Gunn"
date: "2023-04-28"
output: html_document
---

# Project: Effects of admixture on fitness in Neosho Bass populations 
<font size="+1">We assessed the effect of admixture on fitness in two stream populations within the native range of the Neosho Bass (<i>M. velox</i>) which are known to have extensively hybridized with Smallmouth Bass (<i>Micropterus dolomieu</i>). Specifically, we used 14 microsatellite loci in a Bayesian analysis of population structure to estimate proportions of interspecific ancestry in individuals collected from Big Sugar Creek and the Elk River in southwestern Missouri (Central Interior Highlands ecoregion (CIH), North America). We used ancestry inference to identify fish as "Pure Neosho Bass", "Pure Smallmouth Bass", or "Admixed". For each group, we measured age and total length and projected individual growth using the standard paramaterization of the von Bertalanffy growth model, comparing average theoretical maximum length among groups. Finally, we used body condition as a proxy of fitness and generated ancestry-fitness correlations of body condition across the global dataset. We ultimately sought to understand the short-term genetic consequences of admixture for Neosho Bass populations in order to better inform management and long-term viability of distinct, economically and ecologically important sportfish species in the CIH.</font>

## Specific Aim: Ancestry fitness correlation analysis
For this aim, we assessed the linear relationship (correlation) between interspecific ancestry (as measured by proportion of non-native SMB ancestry due to introgression) and individual physiological condition (a proxy of individual fitness) to determine the effect of admixture on fitness in fish in Big Sugar Creek and Elk River in the NB native range. Ancestry-fitness correlation may be positive (fitness increases with ancestry; heterosis by relief from inbreeding depression, hybrid vigor, or adaptive introgression), negative (fitness decreases with ancestry; outbreeding depression by breakdown of coadapted gene complexes, or introduction of deleterious recessive alleles from a large population), or neutral (no effect of ancestry on fitness; formation of a stable hybrid zone). 

## Phases of Analysis
### Phase 1: Physiological condition calculation and analysis
### Phase 2: Linear regression of ancestry and condition

### Libraries needed for analysis
```{r, include=FALSE}
library(readxl)
library(tidyverse)
library(cowplot)
library(boot)
library(rstan)
library(emmeans)
library(scales)
```

## PHASE 1: PHYSIOLOGICAL (BODY) CONDITION CALCULATION
In this phase of analysis, we calculate physiological condition for each fish using the raw total length observation data (not including the back-calculated length-at-age data points ascertained in Analysis 4). To calculate condition, we use the Fulton Condition Index (FCI): F(W,TL) = W/(TL)^3; where F(W,TL) is the FCI at mass (g) W and total length (mm) TL. We measured mass (g) and TL (mm) at the time of sample processing, after fish had died and been preserved on ice. We therefore used measures of mass and TL after some loss of weight due to freeze-thawing. 

### STEP 1: Calculate the Fulton Index of physiological (body) condition (FCI).

#### 1a: Read in full raw data and calculate the log of mass and total length; run the Rmd chunk below.
In this step, I read in the fully filtered raw data (without back-calculated length-at-age data points as estimated in Analysis 4), calculate the log10 of mass (mass_dead) and total length (tl_dead), and calculate the differences between total length at capture (tl_alive) and total length at processing (tl_dead). I then test for normal distribution of differences between tl_alive and tl_dead by examining the frequency distribution of difference values.

##### Read in full raw data and calculate log-values: 
```{r}
# Load in full ancestry data curated in Analysis 3 
load("../ancestry_analysis/data/processed_ancestry_data/full_ancestry_data.Rda")

# Apply log10 transformation to mass and total length data and rename revised dataset 
full_ancestry_data <- full_ancestry_data %>%
  mutate(log_mass_dead = log10(mass_dead)) %>%
  mutate(log_tl_dead = log10(tl_dead)) %>% 
  filter(population == "sample")
```

#### 1b: Determine whether the relationship between weight (mass_dead) and total length (tl_dead) is isometric; run the Rmd chunk below.
The FCI assumes that fish growth is isometric, i.e., the length to weight ratio is cubic, such that weight scales with the cube of of total length, or that the slope (b) of the logarithmic form of the weight-length relationship [log(W) = log(a) + blog(L)] is 3, where W is weight, L is total length, and a is an intercept term. To assess whether the data conform to this assumption, I calculate the log of both mass [log(W)] and total length [log(TL_dead)] and calculate the slope of the linear relationship between these variables. I use the total length of the fish at point of processing (dead) to correspond with mass, which was only recorded at time of processing (dead). I then test whether the slope value of the regression is significantly different from 3, indicating whether the length-weight relationship of these fish is isometric. 

##### Model log-linear relationship of weight (mass_dead) and total length (tl_alive):
```{r}
# Run linear model of mass predicted by total length
lw_mod <- lm(log_mass_dead ~ log_tl_dead, 
             data = full_ancestry_data)

# Extract model coefficients
summary(lw_mod)

# Set up model to determine whether slope of the log-linear model (b ~ 3.12) is different from b = 3.
em_mod <- emtrends(lw_mod, 
                   ~1, 
                   var="log_tl_dead")

# Run hypothesis test
test(em_mod, 
     null = 3)
```

<b>Coefficients of log-linear model for log(W) and log(TL)</b>: <br>

<b>slope</b>= 3.10431 <br>
<b>intercept</b>= -5.16658 <br>
<b><i>r</i>-squared</b>= 0.9604 <br>
<b><i>p</i></b>= 0.9604 <br>

<b>Summary of hypothesis test</b>: <br>

<b><i>p</i></b>= 0.0801 <br>

The slope of the log-linear relationship between log(W) and log(TL) is not significantly different from null of 3. There is support that the relationship between W and TL is isometric.

#### 1c: Calculate FCI; run the Rmd code below:
In this step, we calculate the FCI (g/mm3) for each fish using mass at processing (mass_dead) and total length at processing (tl_dead). We also convert condition to units of mg/cm3 (x 1,000,000) to run the model and for plotting.

##### Calculate FCI:
```{r}
# Calculate condition index and drop NA values. Condition is multiplied by 1000 to reduce unit confusion and to make plotting more aesthetic
condition_data <- full_ancestry_data %>%
  mutate(condition = mass_dead/((tl_dead)^3)) %>%
  mutate(condition_mg_cm = 1000000*condition)

# Save condition data for downstream analyses
save(condition_data, file = "data/condition_data.Rda")
```

### ----------------------- END OF PHASE 1: PHYSIOLOGICAL CONDITION CALCULATION AND ANALYSIS ----------------------- ###

### PHASE 2: LINEAR REGRESSION OF ANCESTRY AND CONDITION
In this step, we use a generalized linear model in a Bayesian hierarchical framework to assess the correlation between non-native SMB ancestry and body condition in the NB native range. 

#### STEP 1: Load and prepare data for analysis; run the Rmd chunk below.
In this step, we load the raw condition data calculated in Phase 1, Step 1c above and prepare for the Bayesian generalized linear model analysis.

##### Load and prepare data for ancestry-condition correlation:
```{r}
# Load condition data generated in Phase 1 above
load("data/condition_data.Rda")

# Generate variables and match names to model 
con <- condition_data$condition_mg_cm
smb <- as.numeric(scale(condition_data$smb)) # Scale to compare effect sizes more accurately
stream <- as.numeric(as.factor(condition_data$river))
stream[which(stream == 3)] <- 2 # recode to 1 and 2 instead of 1 and 3
sex <- as.numeric(as.factor(condition_data$sex))

# Define number of samples in dataset
n <- length(con)

# make stream (riv) and sex factors 0 or 1 (to be compatible with) stan formatting
stream <- stream - 1
sex <- sex - 1

# Define data for model
ac_data <- list(con = con, 
                smb = smb, 
                stream = stream, 
                sex = sex,
                N = n)

# Save data for ancestry-fitness correlation model
save(ac_data, file = "data/ac_data.Rda")
```

### STEP 2: Run generalized linear regression model for body condition and SMB ancestry; run the Rmd chunk below.
In this step, we use a Bayesian hierarchical framework to fit a generalized linear model to assess the linear relationship between body condition and proportion of SMB ancestry. We used 4 cores to run 4 Markov chains, with 2,000 warmup and 4,000 iterations.

##### Fit ancestry-condition correlation model:
```{r}
# Load data for ancestry-condition correlation model
load("data/ac_data.Rda")

# Set MCMC specifications
warmup = 2000       
thin = 2
iter = 4000    

# Run ancestry-condition model
ac_fit <- stan(file = "code/ac_model.stan",  
               data = data,
               chains = 4,        
               thin = thin,
               warmup = warmup,          
               iter = iter,           
               cores = 4)

# Save ancestry-correlation model fit output for downstream analyses
save(ac_fit, file = 'data/ac_fit.Rda')
```

### STEP 3: Summarize and plot model output from ancestry-condition correlation analysis.
In this step, we analyze and plot output results from the generalized linear model of the condition and ancestry proportion.

#### 3a: Run diagnostics on model fit; run the Rmd chunk below:

##### Model diagnostics:
```{r}
# Load ancestry-correlation model fit output
load('data/ac_fit.Rda')

# Visually inspect chains
traceplot(ac_fit)

# Extract summary data for model fit  
sampler_params <- get_sampler_params(ac_fit, inc_warmup = TRUE)

# Summarize model fit
summary(do.call(rbind, sampler_params), digits = 2)
```

#### 3b: Summarize mean and 95 credible intervals for all model coefficients; run the Rmd chunk below:

##### Summarize ancestry-condition fit model output:
```{r}
# Load ancestry-correlation model fit output
load('data/ac_fit.Rda')

# Convert model fit data to dataframe
ac_fit <- as.data.frame(ac_fit)

# find proportion of data points above or below zero for each coefficient to assess significance
length(which(ac_fit$beta1<0))/length(ac_fit$beta1)
length(which(ac_fit$beta2>0))/length(ac_fit$beta2) 
length(which(ac_fit$beta3<0))/length(ac_fit$beta3)
length(which(ac_fit$beta4<0))/length(ac_fit$beta4)

# Gather parameters and estimates at each iteration so that there are only two columns
ac_fit <- ac_fit %>%
  gather(beta1:lp__, key = "parameter", value = "estimate") %>%
  mutate(parameter = factor(parameter))

# Summarize parameter estimates and 95% credible intervals
ac_fit %>%
  group_by(parameter) %>%
  summarize(mean = mean(estimate),
            sd = sd(estimate),
            q2.5 = quantile(estimate, probs = 0.025),
            q25 = quantile(estimate, probs = 0.05), 
            q50 = quantile(estimate, probs = 0.50),
            q75 = quantile(estimate, probs = 0.95),
            q97.5 = quantile(estimate, probs = 0.975))
```

<b>Data summary:</b> <br>

<i>P</i><sub>smb<0</sub> = 0.967 <br>
<i>P</i><sub>stream>0</sub> = 0.986 <br>
<i>P</i><sub>sex<0</sub> = 0.730 <br>
<i>P</i><sub>streamxsmb<0</sub> = 0.587 <br>

We found that SMB ancestry explained substantial negative variation in condition, and we found that Elk River fish had significantly higher condition than Big Sugar Creek fish.

#### 3c: Plot 95% credible intervals for each parameter coefficient; run the Rmd chunk below.

##### Plot ancestry-condition beta coefficient 95% credible intervals: `figures/ac_betas.pdf`
```{r}
# Load ancestry-correlation model fit output
load('data/ac_fit.Rda')

# Convert model fit data to dataframe
ac_fit <- as.data.frame(ac_fit)

# Gather parameters and estimates at each iteration so that there are only two columns, for plotting
ac_fit <- ac_fit %>%
  gather(beta1:lp__, key = "parameter", value = "estimate") %>%
  mutate(parameter = factor(parameter))

# Select only parameters needed for plotting (beta1, beta2, beta3, beta4)
betas <- ac_fit %>%
  filter(parameter == "beta1" |
           parameter == "beta2" |
           parameter == "beta3" |
           parameter == "beta4")

# Plot beta coefficient 95% CIs
jpeg("figures/ac_betas.jpeg", width = 2400, height = 1900, res = 500)

ggplot(betas, aes(x = parameter, y = estimate)) +
  geom_boxplot(fill = "grey") +
  theme_set(theme_cowplot(12)) + 
  labs(x = "Parameter", y = "Estimate") +
  scale_y_continuous(breaks = c(2.5, 2, 1.5, 1, 0.5, 0, -0.5, -1, -1.5, -2), limits = c(-2,2)) + 
  geom_hline(yintercept = 0, size = 1, linetype = "longdash", color = "red") +
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(plot.margin = margin(10,10,10,10))

dev.off()
```

#### 3d: Plot predicted condition values across ancestry along with raw (converted) data; run the Rmd chunk below.

##### Plot ancestry-condition correlation prediction: `figures/ac_predict.pdf`
```{r}
# Load ancestry-correlation model fit output
load('data/ac_fit.Rda')

# Load ancestry-condition data
load('data/ac_data.Rda')

# Load raw condition data
load('data/condition_data.Rda')

# Convert model fit data to dataframe
ac_fit <- as.data.frame(ac_fit)

# Get alpha value 
alpha <- ac_fit$alpha

# Get min and max scaled smb from ac data
min(ac_data$smb)
max(ac_data$smb)

# Get sequence of smb values to predict from, and increase by 0.1 - We opted to extend these values beyond the min and max of the scaled values to make sure we encompassed predicted condition values for smb ancestry values of 1 and 0.
smb <- seq(-6, 6, 0.1) 

# Generate empty matrix to hold predicted condition values for the sequence of smb values given above (smb)
predicted_condition <- matrix(NA, 
                              length(ac_fit$beta1), 
                              length(smb)) 

# Fill empty matrix for prediction plot data
for (i in 1:length(smb)){ # fill empty matrix with predicted values using the model
  predicted_condition[,i] <- alpha +
    ac_fit$beta1*smb[i] + 
    ac_fit$beta2 + 
    ac_fit$beta3 + 
    ac_fit$beta4
}

# Get numeric vectors to hold means and confidence interval estimates from prediction
mean_beta1 <- uci_beta1 <- lci_beta1 <- numeric()

# Predict mean and confidence intervals around trendline
for (i in 1:length(smb)){
  mean_beta1[i] <- mean(predicted_condition[,i]) 
  uci_beta1[i] <- quantile(predicted_condition[,i], probs=0.975)
  lci_beta1[i] <- quantile(predicted_condition[,i], probs=0.025)
}

# Convert mean and confidence interval vectors to data frames 
smb <- as.data.frame(smb)
mean_beta1 <- as.data.frame(mean_beta1)
uci_beta1 <- as.data.frame(uci_beta1)
lci_beta1 <- as.data.frame(lci_beta1)

# Change column names
colnames(smb) <- c("smb")
colnames(mean_beta1) <- c("condition")
colnames(uci_beta1) <- c("uci")
colnames(lci_beta1) <- c("lci")

# Column bind predicted mean and confidence interval estimates
predicted_condition <- cbind(smb,
                             mean_beta1,
                             uci_beta1,
                             lci_beta1)

# Unscale SMB % for plotting 
smb_unscaled <- smb*sd(condition_data$smb)+mean(condition_data$smb)

# Change column name
colnames(smb_unscaled) <- c("smb_us")

# Bind unscaled smb data with predicted data
predicted_condition <- cbind(smb_unscaled, predicted_condition)

# Plot von Bertalanffy curve for global data
jpeg("figures/ac_predict.jpeg", width = 2400, height = 1900, res = 500)

ggplot() + 
  geom_point(data = condition_data, aes(x = smb, y = condition_mg_cm), fill = "grey", show.legend = F, size = 3, pch = 21) + 
  geom_ribbon(data = predicted_condition, aes(x = smb_us, ymin = lci, ymax = uci, fill = "grey", alpha = 0.2), show.legend = F) +
  geom_line(data = predicted_condition, aes(x = smb_us, y = condition, color = "black"), size = 1, show.legend = F) +
  theme_set(theme_cowplot(12)) +
  scale_fill_manual(values = c("grey")) +
  scale_color_manual(values = c("black")) +
  xlim(0,1) +
  ylim(7.5, 17.5) +
  scale_x_continuous(name = "Proportion of SMB ancestry (%)", expand = c(0,0), limits = c(0,1)) +
  scale_y_continuous(labels = number_format(accuracy = 0.001), expand = c(0,0)) +
  ylab(expression(Condition~(mg/cm^{"3"}))) + 
  theme(axis.title = element_text(size = 15)) + 
  theme(axis.text = element_text(size = 15)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(plot.margin = margin(15,15,15,15))

dev.off()
```

### ----------------------- END OF PHASE 2: LINEAR REGRESSION OF ANCESTRY AND CONDITION ----------------------- ###

### ----------------------- END OF ANALYSIS 5: ANCESTRY CONDITION CORRELATION ANALYSIS ----------------------- ###

